Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-18 arxiv/proof_same_cov.tex
section D wip

	 One can see that the expected optimization objective also admits a nice bias-variance decomposition.
	Furthermore, its minimizer only depends on the bias term since the variance term is fixed, and the minimizer of the bias term is precisely $A^{\star} {A^{\star}}^{\top}$.

	The next piece of our proof deals with the prediction loss of hard parameter sharing.
	\begin{claim}\label{claim_pred_err}
		In the setting of Theorem \ref{thm_many_tasks},
		let $\hat{a}_i = \hat{A}^{\top} (\hat{A}\hat{A}^{\top})^{+} \hat{A}_i$.
		We have that the prediction loss of $\hat{\beta}_i^{\MTL} := \hat{B} \hat{A}_i$ satisfies that
		\begin{align*}
			\bigabs{L(\hat{\beta}_i^{\MTL}) - L(B^{\star} \hat{a}_i) - \sigma^2 \norm{\hat{a}_i}^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}}
			\le  n^{-1/4} \left(L(B^{\star} \hat{a}_i) + \sigma^2  \cdot\|\hat a_i\|^2\right) .
		\end{align*}
	\end{claim}
	%The proof of Claim \ref{claim_opt_dist}, Claim \ref{lem_exp_opt}, and Claim \ref{claim_pred_err} can be found in Appendix \ref{app_proof_error_same_cov}.
	Provided with these results, we are ready to prove Theorem \ref{thm_many_tasks}.
	\begin{proof}[Proof of Theorem \ref{thm_many_tasks}]
		Using Claim \ref{claim_pred_err}, we get that the prediction loss of $\hat{\beta}_i^{\MTL}$ is equal to  $L(B^{\star}\hat{a}_i)+\sigma^2\norm{\hat{a}_i}^2\cdot \bigtr{\Sigma(X^{\top}X)^{-1}}$ up to a small enough error. Moreover, Claim \ref{claim_opt_dist} gives directly an upper bound on $\|\hat a_i - a_i^\star\|^2$. With this estimate, we can bound the difference
		$$L(B^{\star}\hat{a}_i)+\sigma^2\norm{\hat{a}_i}^2\cdot \bigtr{\Sigma(X^{\top}X)^{-1}} - L(B^{\star} {a}^\star_i)-\sigma^2\norm{{a}^\star_i}^2\cdot \bigtr{\Sigma(X^{\top}X)^{-1}}.$$
		Combined together, our proof is complete.
%		For the latter, we use Claim \ref{claim_opt_dist} to upper bound the difference between $\norm{\hat{a}_i}^2$ and $\norm{a_i^{\star}}^2$.
%		For $L(B^{\star}\hat{a}_i)$, we again use Claim \ref{claim_opt_dist} to upper bound the distance between $\hat{a}_i$ and $a_i^{\star}$.
%		Combined together, our proof is complete.
	\end{proof}

@@

	For equation \eqref{eq_gA_p2}, 
	%using Lemma \ref{largedeviation} in Appendix \ref{app_tool} and the fact that all moments of $\varepsilon_i$ exist by Assumption \eqref{assmAhigh2}, 
	using Corollary \ref{cor_calE}, we obtain the following with high probability:
	\begin{align}
		|\inner{XB^{\star} (U_AU_A^{\top} - \id_{t\times t})}{\cE}| &\le n^\e \cdot \sigma \cdot \bignormFro{XB^{\star} (U_AU_A^{\top} - \id_{t\times t})} \nonumber \\
		&\le n^{\e} \cdot \sigma \cdot \norm{Z} \cdot \bignormFro{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id_{t\times t})} \nonumber \\
		&\lesssim n^{\e} \cdot \sigma \cdot \sqrt{n} \bignorm{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id_{t\times t})} \label{eq_gA_err2}
%		&\le  n^{c_\infty +c} \cdot \bignorm{\Sigma^{1/2} B^{\star} (U_AU_A^{\top} - \id_{t\times t})}^2 + n^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t.
	\end{align}
	In the second step, we use the fact that $X=Z\Sigma^{1/2}$.
	In the third step, we use Fact \ref{fact_minv}(ii) to bound the operator norm of $Z$ by $\OO(\sqrt{n})$, and use $\bignormFro{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id_{t\times t})}\le \sqrt{t} \bignorm{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id_{t\times t})}$ since the matrix $\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id_{t\times t})$ has rank at most $t$.
	By the AM-GM inequality, equation \eqref{eq_gA_err2} is bounded by the right-hand side of \eqref{eq_gA_err}.

	For equation \eqref{eq_gA_p3}, using Corollary \ref{cor_calE},
	%Lemma \ref{largedeviation} and the fact that all moments of $\varepsilon_i$ exist, 
	we obtain that with high probability,
	\begin{align}
		\abs{\normFro{\cE}^2 - \sigma^2 \cdot n \cdot t}=\abs{\tr\left[ \cal E^\top \id_{n\times n}\cal E\right] - \sigma^2 \cdot n \cdot t} &\le n^{c} \cdot \sigma^2 \| \id_{n\times n}\|_F \nonumber\\
		&= n^{1/2+c} \cdot \sigma^2 \le n^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t. \label{eq_gA_err3}
	\end{align}
	For the inner product between $\cW A$ and $\cE$, we have that with high probability,
	\begin{align}
		\bigabs{\inner{\cW A}{\cE} - \sigma^2\cdot p \cdot r } &= \bigabs{\bigtr{\left(\cE^\top U_X U_X^{\top}\cE  - p\sigma^2 \cdot  \id_{t\times t} \right)  U_AU_A^{\top} }} \nonumber  \\
		&\le \bignormFro{U_{A} U_A^{\top}} \cdot \bignorm{\cE^\top U_X U_X^{\top} \cE - p \sigma^2 \cdot  \id_{t\times t}} \nonumber \\
		&\le \sqrt{r} \cdot n^{\e}\cdot \sigma^2\cdot \|U_X U_X^\top\|_F \nonumber \\
		&\le n^{1/2+c} \cdot \sigma^2 \cdot \sqrt{r}  \le  n^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t.\label{eq_gA_err4}
	\end{align}
	Here in the third step, we apply equation \eqref{vcalAE3} to $\bignorm{\cE^\top U_X U_X^{\top} \cE - p \sigma^2 \cdot  \id_{t\times t}}$ and use that $ \bignormFro{U_{A} U_A^{\top}}=\sqrt{r}$ because $U_A$ has rank $r$. In the fourth step, we use $\|U_X U_X^\top\|_F=\sqrt{p}$ because $U_X$ has rank $p$.

	Combining equations \eqref{eq_gA_err1}, \eqref{eq_gA_err2}, \eqref{eq_gA_err3}, and \eqref{eq_gA_err4}, we obtain equation \eqref{eq_gA_err}.

@@

	\paragraph{Part 2:} The second part is the cross term, whose expectation over $\cE$ is zero.
	Let $b = B^{\star} \hat{a}_i - \beta^{(i)}$ for simplicity.
	Using Corollary \ref{cor_calE}, the concentration error can be bounded as
	\begin{align*}
		\bigabs{\inner{\Sigma^{1/2}b}{\Sigma^{1/2}R_i}}
		&= \bigabs{\inner{X (X^{\top}X)^{-1} \Sigma b \hat{a}_i^{\top}}{\cE}} \\
		&\le \sum_{j=1}^t \abs{\hat{a}_i(j)} \cdot \bigabs{\inner{X (X^{\top} X)^{-1} \Sigma b}{\varepsilon^{(j)}}} \\
		&\le \sum_{j=1}^t \abs{\hat{a}_i(j)} \cdot n^{\e}\sigma \bignorm{X (X^{\top}X)^{-1} \Sigma b} \\
		&\le \sqrt{t} \norm{\hat{a}_i} \cdot n^{\e} \sigma \bignormFro{X (X^{\top} X)^{-1} \Sigma b}.
	\end{align*}
%	\begin{align*}
%		\bigabs{\inner{\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)}{\Sigma^{1/2}R_i}} &= \bigabs{\inner{\hat a_i^\top X(X^\top X)^{-1}\Sigma (B^{\star} \hat{a}_i - \beta_i)}{\cal E}} \\
%		&\le n^{c}\sigma \cdot \bignormFro{\hat a_i^\top X(X^\top X)^{-1}\Sigma (B^{\star} \hat{a}_i - \beta_i)}\\
%		&\le n^{c}\sigma \cdot\|\hat a_i\| \cdot  \|\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)\| \cdot \| (Z^\top Z)^{-1} \|^{1/2} \\
%		&\lesssim n^{-1/2+c}\sigma \cdot\|\hat a_i\| \cdot \bignorm{\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)} \\
%		&\le n^{-1/2+c}\sigma^2  \cdot\|\hat a_i\|^2 + n^{-1/2+c} L(B^{\star} \hat{a}_i), 
%	\end{align*}
	In the first step, we plug in the definition of $R_i$ and re-arrange terms.
	In the second step, we use $\hat{a}_i(j)$ to denote the $j$-th coordinate of $\hat{a}_i$.
	In the third step, we use equation \eqref{vcalE}.
	In the last step, we use $\sum_j |\hat a_i(j)|\le \sqrt{t}\|\hat a_i\|$ by Cauchy-Schwarz inequality.
	Finally, we have
	\begin{align*}
		\bignormFro{ X(X^\top X)^{-1}\Sigma b} &=\left[ {b}^\top \Sigma (X^\top X)^{-1}X^\top X(X^\top X)^{-1}\Sigma b \right]^{1/2} \\
		&\le  \|\Sigma^{1/2} b \|  \cdot \left\| \Sigma^{1/2}(X^\top X)^{-1}\Sigma ^{1/2} \right\|^{1/2}  =   \|\Sigma^{1/2} b \|  \cdot \left\| (Z^\top Z)^{-1}\right\|^{1/2} \\
		&\le n^{-1/2} \cdot \norm{\Sigma^{1/2} b}.
	\end{align*}
%	\begin{align*}
%	\bignormFro{\hat a_i^\top X(X^\top X)^{-1}\Sigma (B^{\star} \hat{a}_i - \beta_i)} &=\|\hat a_i\|\cdot \left[ (B^{\star} \hat{a}_i - \beta_i)^\top \Sigma (X^\top X)^{-1}X^\top X(X^\top X)^{-1}\Sigma (B^{\star} \hat{a}_i - \beta_i) \right]^{1/2} \\
%	&\le \|\hat a_i\|\cdot  \|\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)\|  \cdot \left\| \Sigma^{1/2}(X^\top X)^{-1}\Sigma ^{1/2} \right\|^{1/2} \\
%	&= \|\hat a_i\|\cdot  \|\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)\|  \cdot \left\| (Z^\top Z)^{-1}\right\|^{1/2}
%	\end{align*}
	Above, we use $X=Z\Sigma^{1/2}$. In the last step, we use Fact \ref{fact_minv}(ii) to bound the operator norm of $Z^\top Z$ by $\OO(n^{-1})$. %, and apply the AM-GM inequality.
	One can see that the concentration error from this part is upper bounded by the result in Claim \ref{claim_pred_err}.

>>>>

	 One can see that the expected optimization objective also admits a nice bias-variance decomposition.
	Furthermore, its minimum only depends on the bias term since the variance term is fixed, and the minimizer of the bias term is precisely $A^{\star} {A^{\star}}^{\top}$.

	The next piece of our proof deals with the prediction loss of hard parameter sharing.
	\begin{claim}\label{claim_pred_err}
		In the setting of Theorem \ref{thm_many_tasks},
		let $\hat{a}_i = \hat{A}^{\top} (\hat{A}\hat{A}^{\top})^{+} \hat{A}_i$.
		We have that the prediction loss of $\hat{\beta}_i^{\MTL} := \hat{B} \hat{A}_i$ satisfies that
		\begin{align*}
			\bigabs{L(\hat{\beta}_i^{\MTL}) - L(B^{\star} \hat{a}_i) - \sigma^2 \norm{\hat{a}_i}^2 \cdot \bigtr{\Sigma (X^{\top}X)^{-1}}}
			\le  n^{-1/4} \left(L(B^{\star} \hat{a}_i) + \sigma^2  \cdot\|\hat a_i\|^2\right) .
		\end{align*}
	\end{claim}
	%The proof of Claim \ref{claim_opt_dist}, Claim \ref{lem_exp_opt}, and Claim \ref{claim_pred_err} can be found in Appendix \ref{app_proof_error_same_cov}.
	Provided with these results, we are ready to prove Theorem \ref{thm_many_tasks}.
	\begin{proof}[Proof of Theorem \ref{thm_many_tasks}]
		Using Claim \ref{claim_pred_err}, we get that the prediction loss of $\hat{\beta}_i^{\MTL}$ is equal to  $L(B^{\star}\hat{a}_i)+\sigma^2\norm{\hat{a}_i}^2\cdot \bigtr{\Sigma(X^{\top}X)^{-1}}$ up to a multiplicative error of order $n^{-1/4}$.
		%Moreover, Claim \ref{claim_opt_dist} gives directly an upper bound on $\|\hat a_i - a_i^\star\|^2$. With this estimate, we can bound the difference
		%$$L(B^{\star}\hat{a}_i)+\sigma^2\norm{\hat{a}_i}^2\cdot \bigtr{\Sigma(X^{\top}X)^{-1}} - L(B^{\star} {a}^\star_i)-\sigma^2\norm{{a}^\star_i}^2\cdot \bigtr{\Sigma(X^{\top}X)^{-1}}.$$
		%Combined together, our proof is complete.
		For the latter, we use Claim \ref{claim_opt_dist} to upper bound the difference between $\norm{\hat{a}_i}^2$ and $\norm{a_i^{\star}}^2$.
		For $L(B^{\star}\hat{a}_i)$, we again use Claim \ref{claim_opt_dist} to upper bound the distance between $\hat{a}_i$ and $a_i^{\star}$.
		Combined together, we obtain the difference if we replace $\hat{a}_i$ with $a^{\star}_i$ in Claim \ref{claim_pred_err}, and the proof is complete.
	\end{proof}

@@

	For equation \eqref{eq_gA_p2}, 
	%using Lemma \ref{largedeviation} in Appendix \ref{app_tool} and the fact that all moments of $\varepsilon_i$ exist by Assumption \eqref{assmAhigh2}, 
	using Corollary \ref{cor_calE}, we obtain the following with high probability:
	\begin{align}
		|\inner{XB^{\star} (U_AU_A^{\top} - \id_{t\times t})}{\cE}| &\le n^\e \cdot \sigma \cdot \bignormFro{XB^{\star} (U_AU_A^{\top} - \id_{t\times t})} \nonumber \\
		&\le n^{\e} \cdot \sigma \cdot \norm{Z} \cdot \bignormFro{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id_{t\times t})} \nonumber \\
		&\lesssim n^{\e + 1/2} \cdot \sigma \cdot \bignormFro{\Sigma^{1/2} B^{\star}(U_AU_A^{\top} -\id_{t\times t})} \label{eq_gA_err2}
%		&\le  n^{c_\infty +c} \cdot \bignorm{\Sigma^{1/2} B^{\star} (U_AU_A^{\top} - \id_{t\times t})}^2 + n^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t.
	\end{align}
	In the second step, we use the fact that $X=Z\Sigma^{1/2}$.
	In the third step, we use Fact \ref{fact_minv}(ii) to bound the operator norm of $Z$ by $\OO(\sqrt{n})$.
	By the AM-GM inequality, equation \eqref{eq_gA_err2} is bounded by the right-hand side of \eqref{eq_gA_err}.

	For equation \eqref{eq_gA_p3}, using Corollary \ref{cor_calE},
	%Lemma \ref{largedeviation} and the fact that all moments of $\varepsilon_i$ exist, 
	we obtain that with high probability,
	\begin{align}
		\abs{\normFro{\cE}^2 - \sigma^2 \cdot n \cdot t}=\abs{\tr\left[ \cal E^\top \id_{n\times n}\cal E\right] - \sigma^2 \cdot n \cdot t} &\le n^{c} \cdot \sigma^2 \| \id_{n\times n}\|_F
		= n^{1/2+c} \cdot \sigma^2. \label{eq_gA_err3}
	\end{align}
	For the inner product between $\cW A$ and $\cE$, we have that with high probability,
	\begin{align}
		\bigabs{\inner{\cW A}{\cE} - \sigma^2\cdot p \cdot r } &= \bigabs{\bigtr{\left(\cE^\top U_X U_X^{\top}\cE  - p\sigma^2 \cdot  \id_{t\times t} \right)  U_AU_A^{\top} }} \nonumber  \\
		&\le \bignormFro{U_{A} U_A^{\top}} \cdot \bignorm{\cE^\top U_X U_X^{\top} \cE - p \sigma^2 \cdot  \id_{t\times t}} \nonumber \\
		&\le \sqrt{r} \cdot n^{\e}\cdot \sigma^2\cdot \|U_X U_X^\top\|_F \nonumber \\
		&\le \sqrt{r} \cdot n^{1/2+c} \cdot \sigma^2. \label{eq_gA_err4} %  \le  n^{-c_{\infty}} \cdot \sigma^2 \cdot n \cdot t.
	\end{align}
	Here in the third step, we apply equation \eqref{vcalAE3} to $\bignorm{\cE^\top U_X U_X^{\top} \cE - p \sigma^2 \cdot  \id_{t\times t}}$ and use that $ \bignormFro{U_{A} U_A^{\top}}=\sqrt{r}$ because $U_A$ has rank $r$. In the fourth step, we use $\|U_X U_X^\top\|_F=\sqrt{p}$ because $U_X$ has rank $p$.

	Combining the concentration error estimate for all three parts, we obtain equation \eqref{eq_gA_err}.

@@

	\paragraph{Part 2:} The second part is the cross term, whose expectation over $\cE$ is zero.
	Let $b = B^{\star} \hat{a}_i - \beta^{(i)}$ for simplicity.
	Using Corollary \ref{cor_calE}, the concentration error can be bounded as
	\begin{align*}
		\bigabs{\inner{\Sigma^{1/2}b}{\Sigma^{1/2}R_i}}
		&= \bigabs{\inner{X (X^{\top}X)^{-1} \Sigma b \hat{a}_i^{\top}}{\cE}} \\
		&\le \sum_{j=1}^t \abs{\hat{a}_i(j)} \cdot \bigabs{\inner{X (X^{\top} X)^{-1} \Sigma b}{\varepsilon^{(j)}}} \\
		&\le \sum_{j=1}^t \abs{\hat{a}_i(j)} \cdot n^{\e}\sigma \bignorm{X (X^{\top}X)^{-1} \Sigma b} \\
		&\le \sqrt{t} \norm{\hat{a}_i} \cdot n^{\e} \sigma \bignorm{X (X^{\top} X)^{-1} \Sigma b}.
	\end{align*}
%	\begin{align*}
%		\bigabs{\inner{\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)}{\Sigma^{1/2}R_i}} &= \bigabs{\inner{\hat a_i^\top X(X^\top X)^{-1}\Sigma (B^{\star} \hat{a}_i - \beta_i)}{\cal E}} \\
%		&\le n^{c}\sigma \cdot \bignormFro{\hat a_i^\top X(X^\top X)^{-1}\Sigma (B^{\star} \hat{a}_i - \beta_i)}\\
%		&\le n^{c}\sigma \cdot\|\hat a_i\| \cdot  \|\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)\| \cdot \| (Z^\top Z)^{-1} \|^{1/2} \\
%		&\lesssim n^{-1/2+c}\sigma \cdot\|\hat a_i\| \cdot \bignorm{\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)} \\
%		&\le n^{-1/2+c}\sigma^2  \cdot\|\hat a_i\|^2 + n^{-1/2+c} L(B^{\star} \hat{a}_i), 
%	\end{align*}
	In the first step, we plug in the definition of $R_i$ and re-arrange terms.
	In the second step, we use $\hat{a}_i(j)$ to denote the $j$-th coordinate of $\hat{a}_i$.
	In the third step, we use equation \eqref{vcalE}.
	In the last step, we use $\sum_j |\hat a_i(j)|\le \sqrt{t}\|\hat a_i\|$ by Cauchy-Schwarz inequality.
	Finally, we have
	\begin{align*}
		\bignormFro{ X(X^\top X)^{-1}\Sigma b} &=\left[ {b}^\top \Sigma (X^\top X)^{-1}X^\top X(X^\top X)^{-1}\Sigma b \right]^{1/2} \\
		&\le  \|\Sigma^{1/2} b \|  \cdot \left\| \Sigma^{1/2}(X^\top X)^{-1}\Sigma ^{1/2} \right\|^{1/2}  =   \|\Sigma^{1/2} b \|  \cdot \left\| (Z^\top Z)^{-1}\right\|^{1/2} \\
		&\le n^{-1/2} \cdot \norm{\Sigma^{1/2} b}.
	\end{align*}
%	\begin{align*}
%	\bignormFro{\hat a_i^\top X(X^\top X)^{-1}\Sigma (B^{\star} \hat{a}_i - \beta_i)} &=\|\hat a_i\|\cdot \left[ (B^{\star} \hat{a}_i - \beta_i)^\top \Sigma (X^\top X)^{-1}X^\top X(X^\top X)^{-1}\Sigma (B^{\star} \hat{a}_i - \beta_i) \right]^{1/2} \\
%	&\le \|\hat a_i\|\cdot  \|\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)\|  \cdot \left\| \Sigma^{1/2}(X^\top X)^{-1}\Sigma ^{1/2} \right\|^{1/2} \\
%	&= \|\hat a_i\|\cdot  \|\Sigma^{1/2}(B^{\star} \hat{a}_i - \beta_i)\|  \cdot \left\| (Z^\top Z)^{-1}\right\|^{1/2}
%	\end{align*}
	Above, we use $X=Z\Sigma^{1/2}$. In the last step, we use Fact \ref{fact_minv}(ii) to bound the operator norm of $Z^\top Z$ by $\OO(n^{-1})$. %, and apply the AM-GM inequality.
	One can see that the concentration error from this part is upper bounded by the result in Claim \ref{claim_pred_err}.
